---
title: "Effects of climating warming on photosynthesis"
author: "Dongyue Xie"
date: "2018-11-16"
output: workflowr::wflow_html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
```

## Review

Data from Reich *et al.* 2018, Nature. 

2 sites, 6 blocks, 24 plots in total; 11 Species; 2 treatments; 6 measurements.

- Group 11 species into two categories: Broad-leaf and needle-leaf. 

- Regard measurement day as random effects: total 49 measurement days. Day is a sub-partition of site.


## Species

Previous: Phototsynthesis v.s. Water content across 11 species:

```{r,fig.height=10}
library(lme4)
library(MCMCglmm)
datax=read.csv('/Users/dongyue/Documents/phd y1t1/data analysis/archive_edi/Reich2018NaturePaperDataAug2018.csv')
datax=datax[-c(25,26,580),]
colnames(datax)[3]='trt'
colnames(datax)[4]='plot'
colnames(datax)[13]='water'
datax$plot=factor(datax$plot)

species=c("aceru", "acesa", 'betpa', "quema", "queru", 'poptr', 'rhaca', 'abiba', 'picgl', "pinba", "pinst")

#datax=na.omit(datax)
par(mfrow=c(4,3))
for (sp in species) {
  dat=datax[datax$species==sp,]
  amb=lm(Asat~water,data=dat[dat$trt=='ambient',])
  
  warm=lm(Asat~water,data=dat[dat$trt=='warmed',])
  plot(Asat~water,data=dat,main=sp,ylab='Photosynthesis',
       col=ifelse(dat$trt=='ambient',4,2))
  xx=seq(0.05,0.27,length.out = 100)
  lines(amb$model$water,amb$fitted.values,col=4,lwd=2)
  lines(warm$model$water,warm$fitted.values,col=2,lwd=2)
}

```

## Species

Now across 2 species:

```{r,fig.height=6,fig.width=12}
needle=c('abiba','picgl','pinba','pinst')
datax$species2=factor(ifelse(datax$species %in% needle,'needle','broad'))
species2=c('broad','needle')
par(mfrow=c(1,2))

for (sp in species2) {
  dat=datax[datax$species2==sp,]
  amb=lm(Asat~water,data=dat[dat$trt=='ambient',])
  
  warm=lm(Asat~water,data=dat[dat$trt=='warmed',])
  plot(Asat~water,data=dat,main=sp,ylab='Photosynthesis',
       col=ifelse(dat$trt=='ambient',4,2))
  xx=seq(0.05,0.27,length.out = 100)
  lines(amb$model$water,amb$fitted.values,col=4,lwd=2)
  lines(warm$model$water,warm$fitted.values,col=2,lwd=2)
}
```

## Univariate LMM

Photosynthesis as response; species2, treatment, and water content as predictors.

Model selection gives a Simple model: Photosynthesis ~ water+treatment+species2+treatment:species2+treatment:water+Random effects.

```{r}
datas=na.omit(datax)
datas$ci=datas$ci/100
#year:doy as factor
datas$day=factor(datas$year):factor(datas$doy)
#species two categories: fir(abiba), spruce(picgl), pine(pinba,pinst).
#needle=c('abiba','picgl','pinba','pinst')
#datas$species2=factor(ifelse(datas$species %in% needle,'needle','broad'))

#simple model, only includes asat, water, trt, species
mod1=lmer(Asat~trt*species2*water+(1|day)+(1|block)+(1|plot),data=datas,REML=F)
mod1s=lmer(Asat~(trt+species2+water)^2+(1|day)+(1|block)+(1|plot),data=datas,REML = F)
#anova(mod1s,mod1)

#after some model selection...
mod1ss=lmer(Asat~trt+species2+water+trt:species2+trt:water+(1|day)+(1|block)+(1|plot),data=datas)
#drop1(mod1ss,test='Chisq')
ss.mod1ss=summary(mod1ss)
```

Coefficients:

```{r}
knitr::asis_output("\\scriptsize")
round(ss.mod1ss$coefficients,2)
```

Standard deviation of random components:

```{r}
knitr::asis_output("\\scriptsize")
round(c(sqrt(unlist(ss.mod1ss$varcor)),ss.mod1ss$sigma),2)

```

## Species as random effects

```{r,echo=T}
mod2=lmer(Asat~trt*water+(1|day)+(1|block)+(1|plot)+(1|species2),data=datas)
```
```{r}
ss.mod2=summary(mod2)
```

Coefficients:

```{r}
knitr::asis_output("\\scriptsize")
round(ss.mod2$coefficients,2)
```

Standard deviation of random components:

```{r}
knitr::asis_output("\\scriptsize")
round(c(sqrt(unlist(ss.mod2$varcor)),ss.mod2$sigma),2)
```

## Conditional independence

Is photosynthesis independent of treatment given water content? 

- Conditional multual information. $I(X,Y|Z)=E(\log\frac{P(X,Y|Z)}{P(X|Z)P(Y|Z)})$. It is proportional to the log-likelihood ratio test $G^2$ (they differ by a 2n factor, where n is the sample size) and it is related to the deviance of the tested models(Scutari, M. (2009), page 5, <https://arxiv.org/pdf/0908.3817.pdf>, [ci.test](https://www.rdocumentation.org/packages/bnlearn/versions/4.4/topics/ci.test)).

```{r}
library(bnlearn)
knitr::asis_output("\\scriptsize")
ci.test(datas$Asat,datas$trt,datas$water)
```

- Linear model?

```{r}
knitr::asis_output("\\scriptsize")
mm=lmer(Asat~water+trt+(1|day)+(1|block)+(1|plot)+(1|species2),data=datas)
round(drop1(mm,test='Chisq'),2)
```

## Conditional independence

Is leaf diffusive conductance indepednent of treatment given water content? 

- Conditional multual information

```{r}
knitr::asis_output("\\scriptsize")
ci.test(datas$gs,datas$trt,datas$water)
```

- Linear model?

```{r}
knitr::asis_output("\\scriptsize")
mm=lmer(gs~water+trt+(1|day)+(1|block)+(1|plot)+(1|species2),data=datas)
round(drop1(mm,test='Chisq'),2)
```


## Bivariate response model

Respone: Photosynthesis, water content.  Predictors: treatment. Random effects: day, block, plot.

- A bivariated model for each of 2 species. 

For needle-leaf species:

- First fit unstructured cov matrices. Based on $95%$ credible intervals, only covariance between PS and water in day is significant.  


```{r}
prior2=list(R = list(V = diag(2),nu=1),
               G=list(G1=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G2=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G3=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3)
                  
                  ))
mulmod1=MCMCglmm(cbind(Asat,water)~trait*trt,
                      random=~us(trait):day+idh(trait):block+
                          idh(trait):plot,rcov=~idh(trait):units,
                      prior = prior2,
                      family = c("gaussian","gaussian"),
                      data=datas[datas$species2=='needle',],nitt=13000,burnin = 3000,thin = 10,verbose = F)

```

Fixed effects:

```{r}
round(summary(mulmod1)$solutions[,1:3],3)
```

Cov and Cor - day

```{r}
poster.mode=apply(mulmod1$VCV, 2, posterior.mode)
cov.day=matrix(poster.mode[1:4],2,2)
rownames(cov.day)=c('PS','Water')
colnames(cov.day)=c('PS','Water')
round(cov.day,3)
round(cov2cor(cov.day),3)
```

Sd: block, plot, random error

```{r}
round(sqrt(poster.mode[5:6]),2)
round(sqrt(poster.mode[7:8]),3)
round(sqrt(poster.mode[9:10]),2)
```



## Bivariate model

For broad-leaf species:

```{r}
prior2=list(R = list(V = diag(2),nu=1),
               G=list(G1=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G2=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G3=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3)
                  ))
mulmod1=MCMCglmm(cbind(Asat,water)~trait*trt,
                      random=~us(trait):day+idh(trait):block+
                          idh(trait):plot,rcov=~idh(trait):units,
                      prior = prior2,
                      family = c("gaussian","gaussian"),
                      data=datas[datas$species2=='broad',],nitt=13000,burnin = 3000,thin = 10,verbose = F)

```

Fixed effects:

```{r}
round(summary(mulmod1)$solutions[,1:3],2)
```

Cov and Cor - day

```{r}
poster.mode=apply(mulmod1$VCV, 2, posterior.mode)
cov.day=matrix(poster.mode[1:4],2,2)
rownames(cov.day)=c('PS','Water')
colnames(cov.day)=c('PS','Water')
round(cov.day,3)
round(cov2cor(cov.day),3)
```

Sd: block, plot and random error

```{r}
round(sqrt(poster.mode[5:6]),2)
round(sqrt(poster.mode[7:8]),2)
round(sqrt(poster.mode[9:10]),2)
```

## More

Respones: Photosynthesis v.s. tleaf

Needle-leaf species: None of the covariances between two reponses are significant.

```{r}
prior2=list(R = list(V = diag(2),nu=1),
               G=list(G1=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G2=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G3=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3)
                  ))
mulmod2=MCMCglmm(cbind(Asat,tleaf)~trait*trt,
                      random=~idh(trait):day+idh(trait):block+
                          idh(trait):plot,rcov=~idh(trait):units,
                      prior = prior2,
                      family = c("gaussian","gaussian"),
                      data=datas[datas$species2=='needle',],nitt=13000,burnin = 3000,thin = 10,verbose = F)
```

Fixed effects:

```{r}
round(summary(mulmod2)$solutions[,1:3],2)
```

Sd: block, plot and random error

```{r}
poster.mode=apply(mulmod2$VCV, 2, posterior.mode)
round(sqrt(poster.mode[1:2]),2)
round(sqrt(poster.mode[3:4]),2)
round(sqrt(poster.mode[5:6]),2)
round(sqrt(poster.mode[7:8]),2)
```

## More

Broad-leaf species

```{r}
prior2=list(R = list(V = diag(2),nu=1),
               G=list(G1=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G2=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3),
                  G3=list(V=diag(2),nu=1,alpha.mu=rep(0,2),alpha.V=diag(2)*1e3)
                  ))
mulmod2=MCMCglmm(cbind(Asat,tleaf)~trait*trt,
                      random=~us(trait):day+idh(trait):block+
                        idh(trait):plot,rcov=~us(trait):units,
                      prior = prior2,
                      family = c("gaussian","gaussian"),
                      data=datas[datas$species2=='broad',],nitt=13000,burnin = 3000,thin = 10,verbose = F)
```

Fixed effects:

```{r}
round(summary(mulmod2)$solutions[,1:3],2)
```


Cov and Cor - day

```{r}
poster.mode=apply(mulmod2$VCV, 2, posterior.mode)
cov.day=matrix(poster.mode[1:4],2,2)
rownames(cov.day)=c('PS','Water')
colnames(cov.day)=c('PS','Water')
round(cov.day,3)
round(cov2cor(cov.day),3)
```

Cov and Cor - random error

```{r}
cov.err=matrix(poster.mode[9:12],2,2)
rownames(cov.err)=c('PS','Water')
colnames(cov.err)=c('PS','Water')
round(cov.err,3)
round(cov2cor(cov.err),3)
```

Sd: block, plot

```{r}
round(sqrt(poster.mode[5:6]),2)
round(sqrt(poster.mode[7:8]),2)
```

